Deep Reinforcement Learning usage in games playing\

### 2024-11-18 Paper/main.tex
updated  finished paper

\begin{abstract}

\end{abstract}

\begin{IEEEkeywords}
Deep learning, Reinforcement learning, artificial intelligence, video games
\end{IEEEkeywords}

\section{Introduction}
Games have been a large part of our human society and history from the beginning. Evidence was found that early humans used \text{Talus bones} as rudimentary dice as early as 10,000 B.C\cite{10.1007/978-981-10-0575-6_1       }. As human society developed so did the games they played. The first board game found was in the Levant, dating back to around 7,500 years ago \cite{simpson2007earliest} with a couple of rows of holes carved into limestone. from then on board games developed more complexity from the relative simplicity of games as $Mancala$ in the Mediterranean to $Senet$ in Egypt to $Go$ in China, which is the  oldest board game of mental skill in the world that is still being played\cite{shotwell1994game}. Games like $Go$ and $Chess$ (previously $Chaturanga$) captivated the minds of humans for centuries because they are so complex that no one human could ever master them. This is due to the vast number of possible variations in a single game. In chess for example a typical game is around 40 moves, this could lead $10^{120}$ possible positions according to Claude Shannon\cite{shannon1950xxii}.

@@

\subsection{Mutation rate}
\begin{figure}[h!]
	\centering
	\subfigure[]{\includegraphics[width=0.24\textwidth]{0.2 rate.png}} 
	\subfigure[]{\includegraphics[width=0.24\textwidth]{0.5 rate.png}} 
	\subfigure[]{\includegraphics[width=0.24\textwidth]{0.8 rate.png}} 
	\caption{Bias mutation rate and magintude vs average fitness and highest fitness per generation (a) 0.2  (b) 0.5  (c) 0.8}
	\label{fig:mutations}
\end{figure}
In this section the values for bias mutation power and magnitude are changed and the effect is analysed. For the sake of comparison both parameters are set to the same values as shown in figure \ref{fig:mutations}. In this example the curve of the average fitness is somewhat random between all variations. The number generations needed to reach an optimal solution however is very different. With the values of 0.5 reaching the optimal solution at the quickest rate at 11 generations in comparison to the longest run at a mutation rate of 0.2 reaching the optimal solution after 52 generations. It's also important to note that the highest value for mutation came in the middle place with around 35 generations. This proves that -as with other parameters- for each problem there's a range of values that fit it best which need some variations in order to figure out. 
\section{Conclusion}
\bibliographystyle{unsrt} 
\bibliography{refs}

>>>>

\begin{abstract}
Playing games has been a part of the human story since the very beginning. It was shaped by the societies we created and the technologies we discovered. In recent years Artificial Intelligence and Machine Learning algorithms have come a long way in their complexity and the effect they have on our lives. One of the ways these algorithms were -and are- being used is in playing video games. This allows us to develop and test different models and approaches to solve problems in a constrained environment. In this paper a model is developed using Deep Reinforcement Learning to play the game of \textit{flappy bird}. An analysis is also created of the different parameters and their effect on the output of the model. A reflection on the use of DRL in different fields is also conducted at the end of the paper.
\end{abstract}
\begin{IEEEkeywords}
Deep learning, Reinforcement learning, Deep Reinforcement learning, artificial intelligence, video games
\end{IEEEkeywords}

\section{Introduction}
Games have been a large part of our human society and history from the beginning. Evidence was found that early humans used \text{Talus bones} as rudimentary dice as early as 10,000 B.C\cite{10.1007/978-981-10-0575-6_1}. As human society developed so did the games they played. The first board game found was in the Levant, dating back to around 7,500 years ago \cite{simpson2007earliest} with a couple of rows of holes carved into limestone. from then on board games developed more complexity from the relative simplicity of games as $Mancala$ in the Mediterranean to $Senet$ in Egypt to $Go$ in China, which is the  oldest board game of mental skill in the world that is still being played\cite{shotwell1994game}. Games like $Go$ and $Chess$ (previously $Chaturanga$) captivated the minds of humans for centuries because they are so complex that no one human could ever master them. This is due to the vast number of possible variations in a single game. In chess for example a typical game is around 40 moves, this could lead $10^{120}$ possible positions according to Claude Shannon\cite{shannon1950xxii}.

@@

\subsection{Mutation rate}
\begin{figure}[h!]
	\centering
	\subfigure[]{\includegraphics[width=0.24\textwidth]{0.2 rate.png}} 
	\subfigure[]{\includegraphics[width=0.24\textwidth]{0.5 rate.png}} 
	\subfigure[]{\includegraphics[width=0.24\textwidth]{0.8 rate.png}} 
	\caption{Bias mutation rate and magintude vs average fitness and highest fitness per generation (a) 0.2  (b) 0.5  (c) 0.8}
	\label{fig:mutations}
\end{figure}
In this section the values for bias mutation power and magnitude are changed and the effect is analysed. For the sake of comparison both parameters are set to the same values as shown in figure \ref{fig:mutations}. In this example the curve of the average fitness is somewhat random between all variations. The number generations needed to reach an optimal solution however is very different. With the values of 0.5 reaching the optimal solution at the quickest rate at 11 generations in comparison to the longest run at a mutation rate of 0.2 reaching the optimal solution after 52 generations. It's also important to note that the highest value for mutation came in the middle place with around 35 generations. This proves that -as with other parameters- for each problem there's a range of values that fit it best which need some variations in order to figure out. 
\section{Conclusion}
In conclusion, as the concept of DRL is explored throughout the paper, it is shown how powerful it is in solving unique problems in a controlled environment. As the models generated are flexible they can also be trained to play multiple games, each with their own objectives and score systems. A large scale analysis of the approaches required for such feats is analyzed by Shao et al.\cite{shao2019survey}. From there, it is clear how developments of AI in an application such as video games playing can develop into much more important and critical fields. DRL models are already showing great promises in many other fields. DRL models' ability to learn patterns is specially beneficial in the medical field as analyzed by Zhou et al. \cite{drl_med}, where they show how DRL is used in many fields including breast cancer with an accuracy of 98\% \cite{xu2019attention}. This shows that DRL will be an important part of building AI systems that will change the future in no small capacity.
\bibliographystyle{unsrt} 
\bibliography{refs}
